---
title: "Reproducible Research Project 1"
author: "Asad Bin Imtiaz"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(ggplot2)
```

# Tasks

## Code for reading in the dataset and/or processing the data

```{r cars}
setwd("C:/Users/tgdimas1/RStudioProjects/Reproducible Research/Proj1")
data <- read.csv("activity.csv")
summary(data)
```

Exploration:

Sturnture:
```{r}
str(data)
```

Summary:

```{r}
summary(data)
```
Dimensions:
```{r}
dims <- dim(data)
dims
```
Field Names:
```{r}
names(data)
```

Missing Steps %:

```{r}
total_data <- dims[[1]]
missing_steps <- sum(is.na(data$steps))/total_data * 100
missing_steps
```

Missing dates %:

```{r}
total_data <- dims[[1]]
missing_steps <- sum(is.na(data$date))/total_data * 100
missing_steps
```

Missing intervals %:

```{r}
total_data <- dims[[1]]
missing_steps <- sum(is.na(data$interval))/total_data * 100
missing_steps
```
Sample Dates:

```{r}
head(data$date)
```

Sample Intervals:

```{r}
head(data$interval,20)
```

## Histogram of the total number of steps taken each day

Histogram

```{r}
daily_steps <- with(data, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(daily_steps) <- c("Date", "Steps")
hist(daily_steps$Steps, breaks = seq(0,25000, by=2500), main="Histogram of the total number of steps taken each day",xlab = "Total Steps", ylab="Days", col="blue")
```

## Mean and median number of steps taken each day

Mean of number of steps taken each day

```{r}
mean(daily_steps$Steps)
```

Median of number of steps taken each day

```{r }
median(daily_steps$Steps)
```

## Time series plot of the average number of steps taken

Plot of time series plot of the average number of steps taken:

```{r}
avg_daily_steps <- with(data, aggregate(steps, by = list(interval), FUN = mean, na.rm = TRUE))
names(avg_daily_steps) <- c("Interval", "Activity")
plot(avg_daily_steps$Interval, avg_daily_steps$Activity, main="Time series plot of the average number of steps taken", xlab = "Interval", ylab="Avg Activity", col="blue", type = "l")

```

## Time series plot of the average number of steps taken

The 5-minute interval that, on average, contains the maximum number of steps:

```{r}
avg_daily_steps[which.max(avg_daily_steps$Activity), ]$Interval

```

## Code to describe and show a strategy for imputing missing data

There are many missing steps for several days. Number of missing steps:

```{r}
sum(is.na(data$steps))
```
One strategy to fill the missing data would be to fill it with the mean steps for each day. 
The code would be:

```{r}
avg_steps <- avg_daily_steps$Activity[match(data$interval, avg_daily_steps$Interval)]
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), yes = avg_steps, no = data$steps))
sum(is.na(imputed_data$steps))
```

### Histogram of the total number of steps taken each day after missing values are imputed

Histogram of the total number of steps taken each day after missing values are imputed:

```{r}
imputed_daily_steps <- with(imputed_data, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(imputed_daily_steps) <- c("Date", "Steps")
hist(imputed_daily_steps$Steps, breaks = seq(0,25000, by=2500), main="Histogram of the total number of steps taken each day",xlab = "Total Steps", ylab="Days", col="blue")
```

Mean steps taken each day:

```{r}
mean(imputed_daily_steps$Steps)
```

Median of number of steps taken each day

```{r}
median(imputed_daily_steps$Steps)
```
## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

First we add a new field to indicate weekday or weekend

```{r}
imputed_data$date <- as.Date(strptime(imputed_data$date, format="%Y-%m-%d"))
imputed_data$flg_weekday <- ifelse(as.POSIXlt(imputed_data$date)$wday > 5, "Weekday" , "Weekend")
wday_daily_steps <-aggregate(data = imputed_data, steps~interval+flg_weekday, FUN = mean, na.rm = TRUE)
names(wday_daily_steps) <- c("Interval", "Weekday", "Activity")
ggplot(data=wday_daily_steps,aes(y=Activity,x=Interval,color=Weekday))+
  geom_line()+  labs(title = "Average number of steps taken per 5-minute interval across weekdays and weekends", x = "Interval", y = "Ang Steps") +  facet_wrap(~Weekday, ncol = 1, nrow=2)
```
